---
title: "Visualizing the flights data from the satRday Visualizaiton Challenge"
author: "Dániel Berecz"
date: "2017-07-06"
output: html_document
slug: satrday_flights_analysis
categories: []
tags: []
---

The first satRday was held in Budapest the last year. There was a Visualization
Challenge and I decided to participate on it. Unfortunatelly, because of
technical difficulties, I couldn't send my work in before the deadline, but I often come back to this data set, for testing out new things like this blog.

## Setup

I started to get into R around the satRday, and I learned a lot since then, so I
decided I update the a code a bit. First, I split out the data prepration in
seperate files, and I made the setup more dynamic.

```{r setup, eval=FALSE}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(rprojroot,
               knitr)

root <- find_root(is_rstudio_project)

opts_knit$set(root.dir = root)
rm(list = ls())

bud_url <- "http://budapest.satrdays.org/data/BUD%20flights%202007-2012%20v2.xlsx"
airports_url <- "https://raw.githubusercontent.com/jpatokal/openflights/master/data/airports.dat"

raw_path <- "data/raw/"
flights_name <- paste0(raw_path,"flights.xlsx")
airports_name <- paste0(raw_path,"airports.csv")

dir.create("data/raw", recursive = TRUE)

if(!file.exists(flights_name)) {
  download.file(url = bud_url,
                destfile = flights_name,
                method = "libcurl", mode = "wb")
}

if(!file.exists(airports_name)) {
  download.file(url = airports_url,
                destfile = airports_name,
                method = "libcurl", mode = "wb")
}

# removes the data folder and its contents recursivelly
# unlink(x = "data*", recursive = TRUE)
```

```{r eda, eval=FALSE}
pacman::p_load(readxl,
               tidyverse,
               lubridate)

file_list <- dir("data/raw", pattern = ".xlsx", full.names = TRUE)

flights <- read_excel(file_list, sheet = 1)

colnames(flights) <- tolower(gsub(" ", "_", names(flights)))

flights_cleaned <- flights %>%
  mutate(month = month(date)) %>%
  select(-c(region, commercial_flag, date_half_year, date_year_quarter,
            date_year_month, date, destination)) %>%
  filter(!is.na(city))

char_cols <- names(flights_cleaned[, sapply(flights_cleaned, class) == 'character'])

flights_cleaned[sapply(flights_cleaned, is.character)] <-
  lapply(flights_cleaned[sapply(flights_cleaned, is.character)],
         as.factor)
```

