---
title: The one where I could use backticked reference in real code
author: 'Dániel Berecz'
date: '2017-08-20'
slug: the-one-where-i-could-use-backticked-reference-in-real-code
categories:
  - rstats
tags:
  - draft
---

I alway liked that in Haskell you can inflix a function with backticks like this:

```{r, engine='haskell', eval=FALSE}
take 5 [1..]

-- is the same as

5 `take` [1..]
```

You can do a similar thing in R. In R, when you surround for example an
operator, like `[]`, or `+`, you can use it like a normal function. This is
because these operators are also [function calls][purrr_tut].  
Until now I couldn't really use this in real life code, but the other day I had
to scrape wikipedia and it was pretty handy.

```{r, warning=FALSE, error=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(rvest)
})

# Wiki page for European cities
cities_wiki <- "https://en.wikipedia.org/wiki/List_of_European_cities_by_population_within_city_limits"

# rvest function to read html
eu_cities <- read_html(cities_wiki)

# take the table on the site and extract the cities
cities <- eu_cities %>%
  html_nodes("table") %>%
  html_table() %>%
  `[[`(2) %>% 
  `$`(City)

head(cities)
```


[purrr_tut]: https://jennybc.github.io/purrr-tutorial/ls01_map-name-position-shortcuts.html